---
title: "Hands-on Exercise 1"
author: "Li Jianyi"
---


```{r}
pacman::p_load(tidyverse)
```


```{r}
exam_data <- read_csv("data/Exam_data.csv")
```


Histogram: English Scores
Distribution of English Scores
Shows how English scores are distributed among students.
```{r}
#| echo: false
ggplot(exam_data, aes(x = ENGLISH)) + geom_histogram(binwidth = 5, fill = "skyblue", color = "black") + labs(title = "Distribution of English Scores", x = "Score", y = "Count")
```

Histogram: Maths Scores
Distribution of Maths Scores
Indicates the spread of mathematics performance.
```{r}
#| echo: false
ggplot(exam_data, aes(x = MATHS)) + geom_histogram(binwidth = 5, fill = "orange", color = "black") + labs(title = "Distribution of Maths Scores", x = "Score", y = "Count")
```

Histogram: Science Scores
Distribution of Science Scores
Highlights how students are performing in Science.
```{r}
#| echo: false
ggplot(exam_data, aes(x = SCIENCE)) + geom_histogram(binwidth = 5, fill = "lightgreen", color = "black") + labs(title = "Distribution of Science Scores", x = "Score", y = "Count")
```

Boxplots by Gender
Boxplot of English Scores by Gender
Compares distribution of English scores between genders.
```{r}
#| echo: false
ggplot(exam_data, aes(x = GENDER, y = ENGLISH)) + geom_boxplot(fill = "grey") + labs(title = "English Scores by Gender", x = "Gender", y = "English Score")
```

Boxplot of Maths Scores by Gender
```{r}
#| echo: false
ggplot(exam_data, aes(x = GENDER, y = MATHS)) + geom_boxplot(fill = "lightblue") + labs(title = "Maths Scores by Gender", x = "Gender", y = "Maths Score")
```

Boxplot of Science Scores by Gender
```{r}
#| echo: false
ggplot(exam_data, aes(x = GENDER, y = SCIENCE)) + geom_boxplot(fill = "lightpink") + labs(title = "Science Scores by Gender", x = "Gender", y = "Science Score")
```

Scatter Plots
English vs Maths
Reveals the correlation between English and Math scores.
```{r}
#| echo: false
ggplot(exam_data, aes(x = ENGLISH, y = MATHS)) + geom_point() + labs(title = "English vs Maths", x = "English", y = "Maths")
```

Maths vs Science
Checks if high performance in Math relates to Science.
```{r}
#| echo: false
ggplot(exam_data, aes(x = MATHS, y = SCIENCE)) + geom_point(color = "darkgreen") + labs(title = "Maths vs Science", x = "Maths", y = "Science")
```

Bar Chart of Average Scores by Race
Average Scores by Race
Compares mean subject scores across different races.
```{r}
#| echo: false
exam_data_avg <- exam_data %>% group_by(RACE) %>% summarise(
  English = mean(ENGLISH), Maths = mean(MATHS), Science = mean(SCIENCE)
)

exam_data_avg_long <- tidyr::pivot_longer(exam_data_avg, cols = -RACE, names_to = "Subject", values_to = "Average")

ggplot(exam_data_avg_long, aes(x = RACE, y = Average, fill = Subject)) + geom_bar(stat = "identity", position = "dodge") + labs(title = "Average Subject Scores by Race", x = "Race", y = "Average Score")
```
