---
title: "In-class Exercise 5 : <br> [Mini-Challenge 1](https://vast-challenge.github.io/2025/MC1.html)"
description: ""
date: "2025-05-17"
# date-modified: ""
format: 
  html:
    code-fold: true
    code-summary: "Click to view code"
editor: visual
author: "Li Jianyi"
execute:
  eval: true #- Run the code and show the result.
  echo: true #- Show the code in the output.
  warning: false
  freeze: true
---

<br> <br>

## Overview

The three challenges from [VAST 2025](https://vast-challenge.github.io/2025/index.html).

This in-class exercise challenge is [Mini-Challenge 1](https://vast-challenge.github.io/2025/MC1.html).

# Getting Started

In the code chunk below, p_load() of pacman package is used to load the R packages into R environment.

```{r}
#| code-summary: "Click to view code"
pacman::p_load(jsonlite, tidyverse, SmartEDA, tidygraph, ggraph)
```

## Importing Knowledge Graph Data

In the code chunk below, `fromJSON()` of **jsonlite** package is used to import *MC1_graph.json* file into R and save the output object.

```{r}
#| code-summary: "Click to view code"
kg <- fromJSON("data/MC1_graph.json")
```

### Inspect structure

```{r}
#| code-summary: "Click to view code"
str(kg, max.level = 1)
```

### Extract and inspect

```{r}
#| code-summary: "Click to view code"
nodes_tbl <- as_tibble(kg$nodes)
edges_tbl <- as_tibble(kg$links)
```

## Initial EDA

```{r}
#| code-summary: "Click to view code"
ggplot(data = edges_tbl,
       aes(y = `Edge Type`)) +
  geom_bar()
```

```{r}
#| code-summary: "Click to view code"
ggplot(data = nodes_tbl,
       aes(y = `Node Type`)) +
  geom_bar()
```

## Creating Knowledge Graph

### Step 1: Mapping node id to row index

```{r}
#| code-summary: "Click to view code"
id_map <- tibble(id = nodes_tbl$id,
                 index = seq_len(
                   nrow(nodes_tbl)))
```

This ensures each id from node list is mapped to the correct number.

### Step 2: Map source and target IDs to row indices

```{r}
#| code-summary: "Click to view code"
edges_tbl <- edges_tbl %>%
  left_join(id_map, by = c("source" = "id")) %>%
  rename(from = index) %>%
  left_join(id_map, by = c("target" = "id")) %>%
  rename(to = index)
```

The number of observations in edges_tbl should be the same as before running this code chunk.

Before doing leftjoin, there are only 4 variables. AFter doing the leftjoin, there is two additional variables.

### Step 3: Filter out any unmatched

```{r}
#| code-summary: "Click to view code"
edges_tbl <- edges_tbl %>%
  filter(!is.na(from),!is.na(to))
```

This will get rid of any missing values.

### Step 4: Creating the graph

Lastly, `tbl_graph()` is used to create tidygraph's graph object by using the code chunk below.

```{r}
#| code-summary: "Click to view code"
graph <- tbl_graph(nodes = nodes_tbl,
                   edges = edges_tbl,
                   directed = kg$directed)
```

Directed will be plugged from kg table's directed column.

## Visualising the knowledge graph

```{r}
#| code-summary: "Click to view code"
set.seed(1234)
```

This is to ensure reproducibility. \## Visualising the Whole Graph

```{r}
#| code-summary: "Click to view code"
#| eval: false

ggraph(graph, layout = "fr") + 
  geom_edge_link(alpha = 0.3,               # line, alpha is transparency 
                 colour = "gray") + 
  geom_node_point(aes(color = `Node Type`), # point (plot after line so that it doesn't get covered by line)
                  size = 4) +               # size of point  
  geom_node_text(aes(label = name),         # label using name
                 repel = TRUE,              # prevent overlapping names, force words apart
                 size = 2.5) +
  theme_void()

```

### Visualising the sub-graph

In this section, we are interested to create a sub-graph base on *MemberOf* vaue in *Edge Type* column of the *edges* data frame.

#### Step 1: Filter edges to only "Memberof"

```{r}
#| code-summary: "Click to view code"
graph_memberof <- graph %>%
  activate(edges) %>%                     # Focus on edges table
  filter(`Edge Type` == "MemberOf")       # Filter to MemberOf
```

#### Step 2: Extract only connected nodes (i.e., used in these edges)

```{r}
#| code-summary: "Click to view code"
used_nodes_indices <- graph_memberof %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from,to) %>%            # Only selected variables
  unlist() %>%                  # because it is a graph model, not a list
  unique()
```

This is to eliminate orphan nodes.

#### Step 3: Keep only those nodes

```{r}
#| code-summary: "Click to view code"
graph_memberof <- graph_memberof %>%
  activate(nodes) %>% 
  mutate(row_id = row_number()) %>%
  filter(row_id %in% used_nodes_indices) %>%
  select(-row_id)  # optional clean up
```

#### Plot the sub-graph

```{r}
#| code-summary: "Click to view code"
ggraph(graph_memberof,
       layout = "fr") + 
  geom_edge_link(alpha = 0.5,
                 colour = "gray") +
  geom_node_point(aes(color = `Node Type`),
                  size = 1) +
  geom_node_text(aes(label = name),
                 repel = TRUE,
                 size = 2.5) +
  theme_void()
```
